Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2014-07-17 cordillera-cycle.tex
Remove thousands separator (just checked in CCLI).

Each simulation starts from assumed ice-free conditions at 120\,000 years ago
(120\,ka), and runs to the present. Our modelling domain of 1\,500 by
3\,000\,km
encompasses the entire area covered by the Cordilleran ice sheet at the LGM
(Fig.~\ref{fig:locmap}). The simulations were run on two distinct grids, using
a lower horizontal resolution of 10\,km, and a higher horizontal resolution of
6\,km. These computations were performed on 16 to 128 computing cores at the
Swedish National Supercomputing Centre.

@@

Despite such differences in the timing of attained volume maxima and mimima,
all the model runs show relatively consistent patterns of glaciation
(Fig.~\ref{fig:snapshots}). As a result of scaling factors applied to the
different palaeo-temperature proxy records (Table~\ref{tab:records}), modelled
ice sheet geometries during the last glacial maximum (MIS~2;
Fig.~\ref{fig:snapshots}, lower panels) appear highly similar from one
simulation to the next. They include a ca. 1\,500\,km-long central divide
located about 3\,500\,m\ above sea level (a.s.l.) along the spine of the Rocky
Mountains. The volumes of these modelled last glacial maximum ice sheets are
contained within a close range from 8.24 to 8.66\,m of sea-level equivalent
(s.l.e; Table~\ref{tab:extrema}).
A central ice cap persists in all simulations over the Skeena
Mountains during the MIS~3 ice volume minima (Fig.~\ref{fig:snapshots}, middle
panels). However, the size of the Skeena Mountains ice cap, and the dimensions
of the Cordilleran ice sheet during the MIS~4 glaciation, depend sensitively on
the choice of palaeo-temperature record used to drive the model.

@@

During most of the glacial cycle, modelled ice cover is restricted to disjoint
ice caps centred on major mountain ranges of the North American Cordillera
(Fig.~\ref{fig:duration}). A 2\,500\,km-long continuous ice cover spanning from
the Alaska Range in the northeast to the Rocky Mountains in the southwest only
exists\renote{
    I chosed to write about model output in the present form to discern in from
    the reality.}
for at most 29\,kyr in total over the entire last glacial cycle. However,
except for its marine margin and the northern foothills of the Alaska Range,
the maximum extent of the ice sheet is attained for an even more brief period
of a few thousand years (Figs.~\ref{fig:timeseries}--{fig:duration}). This
result illustrates that the maximum extents of the modelled ice sheet during
MIS~4 and 2 were both short-lived and therefore out of balance with
contemporary climate, which confirms previous inferences of a short-lived
Cordilleran ice sheet based on the geological record \citep{Clague.etal.1980,
Stroeven.etal.2010}.

>>>>

Each simulation starts from assumed ice-free conditions at 120000 years ago
(120\,ka), and runs to the present. Our modelling domain of 1500 by 3000\,km
encompasses the entire area covered by the Cordilleran ice sheet at the LGM
(Fig.~\ref{fig:locmap}). The simulations were run on two distinct grids, using
a lower horizontal resolution of 10\,km, and a higher horizontal resolution of
6\,km. These computations were performed on 16 to 128 computing cores at the
Swedish National Supercomputing Centre.

@@

Despite such differences in the timing of attained volume maxima and mimima,
all the model runs show relatively consistent patterns of glaciation
(Fig.~\ref{fig:snapshots}). As a result of scaling factors applied to the
different palaeo-temperature proxy records (Table~\ref{tab:records}), modelled
ice sheet geometries during the last glacial maximum (MIS~2;
Fig.~\ref{fig:snapshots}, lower panels) appear highly similar from one
simulation to the next. They include a ca. 1500\,km-long central divide
located about 3500\,m\ above sea level (a.s.l.) along the spine of the Rocky
Mountains. The volumes of these modelled last glacial maximum ice sheets are
contained within a close range from 8.24 to 8.66\,m of sea-level equivalent
(s.l.e; Table~\ref{tab:extrema}).
A central ice cap persists in all simulations over the Skeena
Mountains during the MIS~3 ice volume minima (Fig.~\ref{fig:snapshots}, middle
panels). However, the size of the Skeena Mountains ice cap, and the dimensions
of the Cordilleran ice sheet during the MIS~4 glaciation, depend sensitively on
the choice of palaeo-temperature record used to drive the model.

@@

During most of the glacial cycle, modelled ice cover is restricted to disjoint
ice caps centred on major mountain ranges of the North American Cordillera
(Fig.~\ref{fig:duration}). A 2500\,km-long continuous ice cover spanning from
the Alaska Range in the northeast to the Rocky Mountains in the southwest only
exists\renote{
    I chosed to write about model output in the present form to discern in from
    the reality.}
for at most 29\,kyr in total over the entire last glacial cycle. However,
except for its marine margin and the northern foothills of the Alaska Range,
the maximum extent of the ice sheet is attained for an even more brief period
of a few thousand years (Figs.~\ref{fig:timeseries}--{fig:duration}). This
result illustrates that the maximum extents of the modelled ice sheet during
MIS~4 and 2 were both short-lived and therefore out of balance with
contemporary climate, which confirms previous inferences of a short-lived
Cordilleran ice sheet based on the geological record \citep{Clague.etal.1980,
Stroeven.etal.2010}.
